### 2018-02-16 manuscript.tex
typos, figures and citations (a bit of everything)

We gathered a subset of the data from a recent publication \cite{buhl_quasimodo_2016}, which was gracefully made publicly available by the authors\citationneeded{zenodo}.
It shows that flies, which normally become arrhythmic in a Light-Light (LL) regime, are rhythmic (with an period greater than 24h\TODO{phrasing}) when the gene NKCCox is expressed in clock neurons.
The authors have recorded the activity of both TIM/NKCCOK (NKCCox is expressed in their clock neurons) and NKCCOK/+ (the appropriate control) in LL for six days.
Prior to LL, the animals were recorded for a baseline three or four days in normal LD, 12h:12h, conditions.
Two repetitions were performed resulting in a total of 58 animals.
The \texttt{metadata.csv} file, which describes exhaustively all individuals, as well as all the associated result files can be downloaded on zenodo\citationneeded{ibid}.

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{metadata} \hlkwb{<-} \hlkwd{link_dam2_metadata}\hlstd{(}\hlstr{"metadata.csv"}\hlstd{,}\hlstr{"."}\hlstd{)}    \hlcom{# linking}
\hlcom{# print(metadata)                                     # check metadata}
\hlstd{dt} \hlkwb{<-} \hlkwd{load_dam2}\hlstd{(metadata)}                             \hlcom{# loading}
\hlkwd{summary}\hlstd{(dt)}                                           \hlcom{# quick summary}
\end{alltt}
\begin{verbatim}
## behavr table with:
##  58	individuals
##  8	metavariables
##  2	variables
##  1.58722e+05	measurements
##  1	key (id)
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{dt[, label} \hlkwb{:=} \hlkwd{interaction}\hlstd{(}\hlnum{1}\hlopt{:}\hlstd{.N, genotype),} \hlkwc{meta}\hlstd{=T]}
\hlkwd{print}\hlstd{(dt)}
\end{alltt}
\end{kframe}
\end{knitrout}

@@

\begin{figure}[!h]
	%\includegraphics[width=1\textwidth]{fig/fig-3.pdf}
	\caption{{\bf Experiment quality control.}
			Tile plot representing the fraction of time spent moving as a colour intensity.
			Each individual is represented by a row and time, on the x-axis, is binned in 30 minutes.
			A: Uncurated raw data.
			B: Data after the curation step. Time was trimmed and data from dead animals removed. 
			Red `\textbf{+}' symbols show animals that were removed from the subsequent analysis as they had less than five complete days in LL.}
	\label{fig:fig-3}
\end{figure}

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlcom{# make a ggplot object with label on the y and moving on the z axis}
\hlstd{fig3A} \hlkwb{<-} \hlkwd{ggetho}\hlstd{(dt,} \hlkwd{aes}\hlstd{(}\hlkwc{y} \hlstd{= label,} \hlkwc{z} \hlstd{= moving))} \hlopt{+}
  \hlcom{# show data as a tile plot}
  \hlcom{# that is z is a pixel whose intensity maps moving}
  \hlkwd{stat_tile_etho}\hlstd{()} \hlopt{+}
  \hlcom{# add layers to draw annotations to show L and D phases}
  \hlcom{# as white and black, respectivelly}
  \hlcom{# the first layer is for the baseline (until t = 0)}
  \hlkwd{stat_ld_annotations}\hlstd{(}\hlkwc{x_limits} \hlstd{=} \hlkwd{c}\hlstd{(dt[,}\hlkwd{min}\hlstd{(t)],} \hlnum{0}\hlstd{))} \hlopt{+}
  \hlcom{# in the 2nd one, we start at 0 and use grey }
  \hlcom{# instead of black as we work in LL}
  \hlkwd{stat_ld_annotations}\hlstd{(}\hlkwc{x_limits} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{0}\hlstd{, dt[,} \hlkwd{max}\hlstd{(t)]),}
                      \hlkwc{ld_colours} \hlstd{=} \hlkwd{c}\hlstd{(}\hlstr{"white"}\hlstd{,} \hlstr{"grey"}\hlstd{))}
\end{alltt}
\end{kframe}
\end{knitrout}

@@

In addition, we can trim our data to have the same number of days across experiments and individuals.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{dt} \hlkwb{<-} \hlstd{dt[t} \hlopt{%between%} \hlkwd{days}\hlstd{(}\hlkwd{c}\hlstd{(}\hlopt{-}\hlnum{2}\hlstd{,} \hlnum{6}\hlstd{))]}
\hlcom{# same as above}
\hlstd{fig3B} \hlkwb{<-} \hlkwd{ggetho}\hlstd{(dt,} \hlkwd{aes}\hlstd{(}\hlkwc{y} \hlstd{= label,} \hlkwc{z} \hlstd{= moving))} \hlopt{+}
    \hlkwd{stat_tile_etho}\hlstd{()} \hlopt{+}
    \hlkwd{stat_ld_annotations}\hlstd{(}\hlkwc{x_limits} \hlstd{=} \hlkwd{c}\hlstd{(dt[,} \hlkwd{min}\hlstd{(t)],} \hlnum{0}\hlstd{))} \hlopt{+}
    \hlkwd{stat_ld_annotations}\hlstd{(}\hlkwc{x_limits} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{0}\hlstd{, dt[,} \hlkwd{max}\hlstd{(t)]),}
                        \hlkwc{ld_colours} \hlstd{=} \hlkwd{c}\hlstd{(}\hlstr{"white"}\hlstd{,} \hlstr{"grey"}\hlstd{))}
\end{alltt}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlcom{# we also show a subset of this figure in 4A}
\hlstd{figS1A} \hlkwb{<-} \hlkwd{ggetho}\hlstd{(dt,} \hlkwd{aes}\hlstd{(}\hlkwc{z} \hlstd{= moving),} \hlkwc{multiplot} \hlstd{=} \hlnum{2}\hlstd{)} \hlopt{+}
            \hlcom{# bars n the z axis could }
            \hlcom{#  one could also use stat_tile_etho}
            \hlkwd{stat_bar_tile_etho}\hlstd{()} \hlopt{+}
            \hlcom{# divide plot by individual}
            \hlkwd{facet_wrap}\hlstd{(} \hlopt{~} \hlstd{label,} \hlkwc{ncol} \hlstd{=} \hlnum{4}\hlstd{)} \hlopt{+}
            \hlcom{# rename the y axis}
            \hlkwd{scale_y_discrete}\hlstd{(}\hlkwc{name}\hlstd{=}\hlstr{"Day"}\hlstd{)}
\end{alltt}
\end{kframe}
\end{knitrout}

\subsection*{Periodograms}
Ultimately, in order to quantify periodicity and rhythmicity, we compute periodograms.
Several methods are implemented in \texttt{zeitbebr} ($\chi{}^2$, Lomb-Scargle and autocorrelation).
In this example, we generate $\chi{}^2$ periodograms and lay them out in a grid.
Periodograms dor the subset of eight animals used in Fig~\ref{fig:fig-4}A is shown in Fig~\ref{fig:fig-4}B. 
See \nameref{S1-Fig}B for the visualisation of all individuals.

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{dt_ll} \hlkwb{<-} \hlstd{dt[t} \hlopt{>} \hlkwd{days}\hlstd{(}\hlnum{1}\hlstd{)]}
\hlstd{per_dt} \hlkwb{<-} \hlkwd{periodogram}\hlstd{(moving,}
                        \hlstd{dt_ll,}
                        \hlkwc{resample_rate} \hlstd{=} \hlnum{1}\hlopt{/}\hlkwd{mins}\hlstd{(}\hlnum{10}\hlstd{),}
                        \hlkwc{FUN}\hlstd{=chi_sq_periodogram)}

\hlstd{per_dt} \hlkwb{<-} \hlkwd{find_peaks}\hlstd{(per_dt)}
\hlcom{# we also show a subset of this figure in supplementary materials}
\hlstd{figS1B} \hlkwb{<-} \hlkwd{ggperio}\hlstd{(per_dt,} \hlkwd{aes}\hlstd{(}\hlkwc{y} \hlstd{= power,} \hlkwc{peak}\hlstd{=peak))} \hlopt{+}
                  \hlcom{# periododogram drawn as a line}
                  \hlkwd{geom_line}\hlstd{()} \hlopt{+}
                  \hlcom{# the significance line in red}
                  \hlkwd{geom_line}\hlstd{(}\hlkwd{aes}\hlstd{(}\hlkwc{y} \hlstd{= signif_threshold),} \hlkwc{colour}\hlstd{=}\hlstr{"red"}\hlstd{)} \hlopt{+}
                  \hlcom{# point and text at the peak}
                  \hlkwd{geom_peak}\hlstd{()} \hlopt{+}
                  \hlcom{# divide plot by individual}
                  \hlkwd{facet_wrap}\hlstd{(} \hlopt{~} \hlstd{label,} \hlkwc{ncol} \hlstd{=} \hlnum{4}\hlstd{)}
\end{alltt}
\end{kframe}
\end{knitrout}

\begin{figure}[!h]
	%\includegraphics[width=1\textwidth]{fig/fig-4.pdf}
	\caption{{\bf Visualisation of the periodicity in activity of eight representative animals.}
		A: Double plotted actograms showing activity over the experiment. Transition from LD to LL happens at day 0.
		B: $\chi{}^2$ periodograms over the LL part of the experiment matching the animals in A.
		The blue cross represents the first peak (if present) above the significance threshold (red line).
		Titles on top of each facet refer to the label allocated to each individual.
		See \nameref{S1-Fig} for all 52 animals.
	}
	\label{fig:fig-4}
\end{figure}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{fig5A} \hlkwb{<-} \hlkwd{ggperio}\hlstd{(per_dt,} \hlkwd{aes}\hlstd{(}\hlkwc{y} \hlstd{= power} \hlopt{-} \hlstd{signif_threshold,}
                             \hlkwc{colour} \hlstd{= genotype))} \hlopt{+}
          \hlcom{# periododogram shown as a line for population mean}
          \hlcom{# and bootstrap error bars}
          \hlkwd{stat_pop_etho}\hlstd{(}\hlkwc{method} \hlstd{= ggplot2}\hlopt{::}\hlstd{mean_cl_boot)} \hlopt{+}
          \hlcom{# rename x and y axis }
          \hlkwd{scale_y_continuous}\hlstd{(}\hlkwc{name}\hlstd{=}\hlstr{"Relative power"}\hlstd{)} \hlopt{+}
          \hlkwd{scale_x_hours}\hlstd{(}\hlstr{"Period"}\hlstd{)}
\end{alltt}
\end{kframe}
\end{knitrout}

\begin{figure}[!h]
	%\includegraphics[width=1\textwidth]{fig/fig-5.pdf}
	\caption{{\bf Population statistics on circadian phenotype.}
			A: Average periodograms. 
			      The aggregated relative power of the periodogram of all animals.
			      The solid lines and the shaded areas show population means and their 95\% bootstrap confidence interval, respectively.
			B: Frequencies of rhythmic animals.
			      Number of rhythmic animals (\emph{i.e.} with a significant peak) in each genotypes.
			      Dark and clear fillings indicate rhythmic and arhythmic animals, respectively.
      C: Peak periodicity power and average.
			      Values of the peak period for animals with a significant peak (\emph{i.e.} rhythmic).
			      Individual animals are shown by dots whose size represent relative power of the peak period.
			      The error bars are 95\% bootstrap confidence interval on the population mean.
			      }
	\label{fig:fig-5}
\end{figure}

To further quantify this difference, we can show the number of rhythmic animals -- \emph{i.e.} individuals for which a peak was found -- in each groups (see Fig~\ref{fig:fig-5}B).
Then, we can compare the average value of the peak for the rhythmic animals (see Fig~\ref{fig:fig-5}C).
First of all, we compute a summary per individual (\texttt{by=id}):

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{summary_dt} \hlkwb{<-} \hlstd{per_dt[,}\hlkwd{.}\hlstd{(}
                        \hlkwc{first_peak_period} \hlstd{= period[peak}\hlopt{==}\hlnum{1}\hlstd{],}
                        \hlkwc{first_peak_rel_power} \hlstd{= power[peak}\hlopt{==}\hlnum{1}\hlstd{]} \hlopt{-} \hlstd{signif_threshold[peak}\hlopt{==}\hlnum{1}\hlstd{],}
                        \hlkwc{is_rhythmic} \hlstd{=} \hlkwd{any}\hlstd{(peak}\hlopt{==}\hlnum{1}\hlstd{)),}
                      \hlkwc{by}\hlstd{=id]}
\hlcom{# rejoin metadata}
\hlstd{summary_dt} \hlkwb{<-} \hlkwd{rejoin}\hlstd{(summary_dt)}
\end{alltt}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlcom{# standard ggplot}
\hlstd{fig5C} \hlkwb{<-} \hlkwd{ggplot}\hlstd{(summary_dt,} \hlkwd{aes}\hlstd{(}\hlkwc{y} \hlstd{= first_peak_period,}
                                \hlkwc{x} \hlstd{= genotype))} \hlopt{+}
              \hlcom{# draw the mean of each genotype group}
              \hlkwd{stat_summary}\hlstd{(}\hlkwc{fun.y} \hlstd{= mean,} \hlkwc{geom} \hlstd{=} \hlstr{"point"}\hlstd{,} \hlkwc{shape}\hlstd{=}\hlnum{3}\hlstd{)} \hlopt{+}
              \hlcom{# draw bootstrap confidence intervals}
              \hlkwd{stat_summary}\hlstd{(}\hlkwc{fun.data} \hlstd{= mean_cl_boot,} \hlkwc{geom} \hlstd{=} \hlstr{"errorbar"}\hlstd{)} \hlopt{+}
              \hlcom{# shows all individuals as points. }
              \hlcom{# the size of the point expresses the power of the peak}
              \hlkwd{geom_jitter}\hlstd{(}\hlkwd{aes}\hlstd{(}\hlkwc{colour}\hlstd{= genotype,}
                              \hlkwc{size}\hlstd{=first_peak_rel_power),}
                              \hlkwc{alpha}\hlstd{=}\hlnum{.67}\hlstd{)} \hlopt{+}
              \hlcom{# We would like to convert time in hour}
              \hlkwd{scale_y_hours}\hlstd{(}\hlstr{"Period"}\hlstd{)}
\end{alltt}
\end{kframe}
\end{knitrout}

\texttt{R} provides one of the richest statistical toolbox available, which allows users to go deeper tin he analysis of the extracted variables.
One could, for instance, perform a $\chi{}^2$ test on the number of rhythmic \emph{vs} arhythmic flies in both genotypes.
To address the same question, we fit a binomial generalised linear model:

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{fit} \hlkwb{<-} \hlkwd{glm}\hlstd{(is_rhythmic} \hlopt{~} \hlstd{genotype, summary_dt,} \hlkwc{family}\hlstd{=}\hlstr{"binomial"}\hlstd{)}

>>>>

We gathered a subset of the data from a recent publication \cite{buhl_quasimodo_2016}, which was gracefully made publicly available by the authors\cite{ogueta_ll_2018}.
It shows that flies, which normally become arrhythmic in a Light-Light (LL) regime, are rhythmic (with an period greater than 24h\TODO{phrasing}) when the gene NKCCox is expressed in clock neurons.
The authors have recorded the activity of both TIM/NKCCOK (NKCCox is expressed in their clock neurons) and NKCCOK/+ (the appropriate control) in LL for six days.
Prior to LL, the animals were recorded for a baseline three or four days in normal LD, 12h:12h, conditions.
Two repetitions were performed resulting in a total of 58 animals.
The \texttt{metadata.csv} file, which describes exhaustively all individuals, as well as all the associated result files can be downloaded at \href{https://zenodo.org/record/1172980}{https://zenodo.org/record/1172980}.

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{metadata} \hlkwb{<-} \hlkwd{link_dam2_metadata}\hlstd{(}\hlstr{"metadata.csv"}\hlstd{,} \hlstr{"."}\hlstd{)}   \hlcom{# linking}
\hlcom{# print(metadata)                                     # check metadata}
\hlstd{dt} \hlkwb{<-} \hlkwd{load_dam2}\hlstd{(metadata)}                             \hlcom{# loading}
\hlkwd{summary}\hlstd{(dt)}                                           \hlcom{# quick summary}
\end{alltt}
\begin{verbatim}
## behavr table with:
##  58	individuals
##  8	metavariables
##  2	variables
##  1.58722e+05	measurements
##  1	key (id)
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{dt[, label} \hlkwb{:=} \hlkwd{interaction}\hlstd{(}\hlnum{1}\hlopt{:}\hlstd{.N, genotype),} \hlkwc{meta} \hlstd{= T]}
\hlkwd{print}\hlstd{(dt)}
\end{alltt}
\end{kframe}
\end{knitrout}

@@

\begin{figure}[!h]
%    \includegraphics[width=1\textwidth]{fig/fig-3.pdf}
	\caption{{\bf Experiment quality control.}
			Tile plot showing the fraction of time spent moving as a colour intensity.
			Each individual is represented by a row and time, on the x-axis, is binned in 30 minutes.
			A: Uncurated raw data.
			B: Data after the curation step. Time was trimmed and data from dead animals removed. 
			Red `\textbf{+}' symbols show animals that were removed from the subsequent analysis as they had less than five complete days in LL.}
	\label{fig:fig-3}
\end{figure}

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlcom{# make a ggplot object with label on the y and moving on the z axis}
\hlstd{fig3A} \hlkwb{<-} \hlkwd{ggetho}\hlstd{(dt,} \hlkwd{aes}\hlstd{(}\hlkwc{y} \hlstd{= label,} \hlkwc{z} \hlstd{= moving))} \hlopt{+}
  \hlcom{# show data as a tile plot}
  \hlcom{# that is z is a pixel whose intensity maps moving}
  \hlkwd{stat_tile_etho}\hlstd{()} \hlopt{+}
  \hlcom{# add layers to draw annotations to show L and D phases}
  \hlcom{# as white and black, respectively}
  \hlcom{# the first layer is for the baseline (until t = 0)}
  \hlkwd{stat_ld_annotations}\hlstd{(}\hlkwc{x_limits} \hlstd{=} \hlkwd{c}\hlstd{(dt[,}\hlkwd{min}\hlstd{(t)],} \hlnum{0}\hlstd{))} \hlopt{+}
  \hlcom{# in the 2nd one, we start at 0 and use grey }
  \hlcom{# instead of black as we work in LL}
  \hlkwd{stat_ld_annotations}\hlstd{(}\hlkwc{x_limits} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{0}\hlstd{, dt[,} \hlkwd{max}\hlstd{(t)]),}
                      \hlkwc{ld_colours} \hlstd{=} \hlkwd{c}\hlstd{(}\hlstr{"white"}\hlstd{,} \hlstr{"grey"}\hlstd{))}
\end{alltt}
\end{kframe}
\end{knitrout}

@@

In addition, we can trim our data to have the same number of days across experiments and individuals.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlcom{# filter dt between -2d and 6d}
\hlstd{dt} \hlkwb{<-} \hlstd{dt[t} \hlopt{%between%} \hlkwd{days}\hlstd{(}\hlkwd{c}\hlstd{(}\hlopt{-}\hlnum{2}\hlstd{,} \hlnum{6}\hlstd{))]}
\hlcom{# same as above}
\hlstd{fig3B} \hlkwb{<-} \hlkwd{ggetho}\hlstd{(dt,} \hlkwd{aes}\hlstd{(}\hlkwc{y} \hlstd{= label,} \hlkwc{z} \hlstd{= moving))} \hlopt{+}
    \hlkwd{stat_tile_etho}\hlstd{()} \hlopt{+}
    \hlkwd{stat_ld_annotations}\hlstd{(}\hlkwc{x_limits} \hlstd{=} \hlkwd{c}\hlstd{(dt[,} \hlkwd{min}\hlstd{(t)],} \hlnum{0}\hlstd{))} \hlopt{+}
    \hlkwd{stat_ld_annotations}\hlstd{(}\hlkwc{x_limits} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{0}\hlstd{, dt[,} \hlkwd{max}\hlstd{(t)]),}
                        \hlkwc{ld_colours} \hlstd{=} \hlkwd{c}\hlstd{(}\hlstr{"white"}\hlstd{,} \hlstr{"grey"}\hlstd{))}
\end{alltt}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlcom{# we also show a subset of this figure in 4A}
\hlstd{figS1A} \hlkwb{<-} \hlkwd{ggetho}\hlstd{(dt,} \hlkwd{aes}\hlstd{(}\hlkwc{z} \hlstd{= moving),} \hlkwc{multiplot} \hlstd{=} \hlnum{2}\hlstd{)} \hlopt{+}
            \hlcom{# bars n the z axis could }
            \hlcom{# one could also use stat_tile_etho}
            \hlkwd{stat_bar_tile_etho}\hlstd{()} \hlopt{+}
            \hlcom{# split plot by individual}
            \hlkwd{facet_wrap}\hlstd{(} \hlopt{~} \hlstd{label,} \hlkwc{ncol} \hlstd{=} \hlnum{4}\hlstd{)} \hlopt{+}
            \hlcom{# rename the y axis}
            \hlkwd{scale_y_discrete}\hlstd{(}\hlkwc{name} \hlstd{=} \hlstr{"Day"}\hlstd{)}
\end{alltt}
\end{kframe}
\end{knitrout}

\subsection*{Periodograms}
Ultimately, in order to quantify periodicity and rhythmicity, we compute periodograms.
Several methods are implemented in \texttt{zeitbebr} ($\chi{}^2$, Lomb-Scargle and autocorrelation).
In this example, we generate $\chi{}^2$ periodograms and lay them out in a grid.
Periodograms for the subset of eight animals used in Fig~\ref{fig:fig-4}A is shown in Fig~\ref{fig:fig-4}B. 
See \nameref{S1-Fig}B for the visualisation of all individuals.

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlcom{# only the LL data}
\hlstd{dt_ll} \hlkwb{<-} \hlstd{dt[t} \hlopt{>} \hlkwd{days}\hlstd{(}\hlnum{1}\hlstd{)]}
\hlcom{# compute chi sqr periodogram }
\hlstd{per_dt} \hlkwb{<-} \hlkwd{periodogram}\hlstd{(moving,}
                        \hlstd{dt_ll,}
                        \hlkwc{resample_rate} \hlstd{=} \hlnum{1} \hlopt{/} \hlkwd{mins}\hlstd{(}\hlnum{10}\hlstd{),}
                        \hlkwc{FUN}\hlstd{=chi_sq_periodogram)}

\hlstd{per_dt} \hlkwb{<-} \hlkwd{find_peaks}\hlstd{(per_dt)}
\hlcom{# we also show a subset of this figure in supplementary materials}
\hlstd{figS1B} \hlkwb{<-} \hlkwd{ggperio}\hlstd{(per_dt,} \hlkwd{aes}\hlstd{(}\hlkwc{y} \hlstd{= power,} \hlkwc{peak} \hlstd{= peak))} \hlopt{+}
                  \hlcom{# periododogram drawn as a line}
                  \hlkwd{geom_line}\hlstd{()} \hlopt{+}
                  \hlcom{# the significance line in red}
                  \hlkwd{geom_line}\hlstd{(}\hlkwd{aes}\hlstd{(}\hlkwc{y} \hlstd{= signif_threshold),} \hlkwc{colour} \hlstd{=} \hlstr{"red"}\hlstd{)} \hlopt{+}
                  \hlcom{# point and text at the peak}
                  \hlkwd{geom_peak}\hlstd{()} \hlopt{+}
                  \hlcom{# divide plot by individual}
                  \hlkwd{facet_wrap}\hlstd{(} \hlopt{~} \hlstd{label,} \hlkwc{ncol} \hlstd{=} \hlnum{4}\hlstd{)}
\end{alltt}
\end{kframe}
\end{knitrout}

\begin{figure}[!h]
%	\includegraphics[width=1\textwidth]{fig/fig-4.pdf}
	\caption{{\bf Visualisation of the periodicity in activity of eight representative animals.}
		A: Double-plotted actograms showing activity over the experiment. Transition from LD to LL happens at day 0.
		B: $\chi{}^2$ periodograms over the LL part of the experiment matching the animals in A.
		The blue cross represents the first peak (if present) above the significance threshold (red line).
		Titles on top of each facet refer to the label allocated to each individual.
		See \nameref{S1-Fig} for all 52 animals.
	}
	\label{fig:fig-4}
\end{figure}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlcom{# display periodogram }
\hlstd{fig5A} \hlkwb{<-} \hlkwd{ggperio}\hlstd{(per_dt,} \hlkwd{aes}\hlstd{(}\hlkwc{y} \hlstd{= power} \hlopt{-} \hlstd{signif_threshold,}
                             \hlkwc{colour} \hlstd{= genotype))} \hlopt{+}
          \hlcom{# periododogram shown as a line for population mean}
          \hlcom{# and bootstrap error bars}
          \hlkwd{stat_pop_etho}\hlstd{(}\hlkwc{method} \hlstd{= ggplot2}\hlopt{::}\hlstd{mean_cl_boot)} \hlopt{+}
          \hlcom{# rename x and y axis }
          \hlkwd{scale_y_continuous}\hlstd{(}\hlkwc{name} \hlstd{=} \hlstr{"Relative power"}\hlstd{)} \hlopt{+}
          \hlkwd{scale_x_hours}\hlstd{(}\hlstr{"Period"}\hlstd{)}
\end{alltt}
\end{kframe}
\end{knitrout}

\begin{figure}[!h]
%  	\includegraphics[width=1\textwidth]{fig/fig-5.pdf}
	\caption{{\bf Population statistics on circadian phenotype.}
			A: Average periodograms. 
			      The aggregated relative power of the periodogram of all animals.
			      The solid lines and the shaded areas show population means and their 95\% bootstrap confidence interval, respectively.
			B: Frequencies of rhythmic animals.
			      Number of rhythmic animals (\emph{i.e.} with a significant peak) in each genotypes.
			      Dark and clear fillings indicate rhythmic and arhythmic animals, respectively.
      C: Peak periodicity power and average.
			      Values of the peak period for animals with a significant peak (\emph{i.e.} rhythmic).
			      Individual animals are shown by dots whose size represent relative power of the peak period.
			      The error bars are 95\% bootstrap confidence interval on the population mean.
			      }
	\label{fig:fig-5}
\end{figure}

To further quantify this difference, we can show the number of rhythmic animals -- \emph{i.e.} individuals for which a peak was found -- in each group (see Fig~\ref{fig:fig-5}B).
Then, we can compare the average value of the peak for the rhythmic animals (see Fig~\ref{fig:fig-5}C).
First of all, we compute a summary per individual (\texttt{by=id}):

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{summary_dt} \hlkwb{<-}
    \hlstd{per_dt[,}
          \hlkwd{.}\hlstd{(}
           \hlkwc{first_peak_period} \hlstd{= period[peak} \hlopt{==} \hlnum{1}\hlstd{],}
           \hlstd{\{} \hlcom{# \{\} can be used for tmp variables}
             \hlstd{signif} \hlkwb{=} \hlstd{signif_threshold[peak} \hlopt{==} \hlnum{1}\hlstd{]}
             \hlstd{power} \hlkwb{=} \hlstd{power[peak} \hlopt{==} \hlnum{1}\hlstd{]}
             \hlstd{first_peak_rel_power} \hlkwb{=}  \hlstd{power} \hlopt{-} \hlstd{signif}
            \hlstd{\},}
           \hlkwc{is_rhythmic} \hlstd{=} \hlkwd{any}\hlstd{(peak} \hlopt{==} \hlnum{1}\hlstd{)}
           \hlstd{),}
           \hlkwc{by}\hlstd{=id]}

\hlcom{# rejoin metadata}
\hlstd{summary_dt} \hlkwb{<-} \hlkwd{rejoin}\hlstd{(summary_dt)}
\end{alltt}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlcom{# standard ggplot}
\hlstd{fig5C} \hlkwb{<-} \hlkwd{ggplot}\hlstd{(summary_dt,} \hlkwd{aes}\hlstd{(}\hlkwc{y} \hlstd{= first_peak_period,}
                                \hlkwc{x} \hlstd{= genotype))} \hlopt{+}
              \hlcom{# draw the mean of each genotype group}
              \hlkwd{stat_summary}\hlstd{(}\hlkwc{fun.y} \hlstd{= mean,} \hlkwc{geom} \hlstd{=} \hlstr{"point"}\hlstd{,} \hlkwc{shape}\hlstd{=}\hlnum{3}\hlstd{)} \hlopt{+}
              \hlcom{# draw bootstrap confidence intervals}
              \hlkwd{stat_summary}\hlstd{(}\hlkwc{fun.data} \hlstd{= mean_cl_boot,} \hlkwc{geom} \hlstd{=} \hlstr{"errorbar"}\hlstd{)} \hlopt{+}
              \hlcom{# shows all individuals as points}
              \hlcom{# the size of the point expresses the power of the peak}
              \hlkwd{geom_jitter}\hlstd{(}\hlkwd{aes}\hlstd{(}\hlkwc{colour} \hlstd{= genotype,}
                              \hlkwc{size} \hlstd{= first_peak_rel_power),}
                              \hlkwc{alpha} \hlstd{=} \hlnum{0.67}\hlstd{)} \hlopt{+}
              \hlcom{# We would like to convert time in hour}
              \hlkwd{scale_y_hours}\hlstd{(}\hlstr{"Period"}\hlstd{)}
\end{alltt}
\end{kframe}
\end{knitrout}

\texttt{R} provides one of the richest statistical toolbox available, which allows users to go deeper in the analysis of the extracted variables.
One could, for instance, perform a $\chi{}^2$ test on the number of rhythmic \emph{vs} arhythmic flies in both genotypes.
To address the same question, we fit a binomial generalised linear model:

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{fit} \hlkwb{<-} \hlkwd{glm}\hlstd{(is_rhythmic} \hlopt{~} \hlstd{genotype, summary_dt,} \hlkwc{family} \hlstd{=} \hlstr{"binomial"}\hlstd{)}
